The Shape and Color of Politics

### 2023-10-09 chapter_2/src/slides/slides.qmd
a few more edits

```{r}
#| label: setup-block
# Set seed
set.seed(12102023)
# Load libraries
box::use(
  ggplot2[
    ggplot
    , aes
    , geom_density
    , theme_minimal
    , geom_vline
  ]
  , modelsummary[
    datasummary_skim
    , datasummary_balance
    , modelsummary
  ]
  , ../R/predicted_prob_bar[...]
  , data.table[...]
)

# Load cleaned data
list_df <- list()
list_df[["cleaned"]] <- read.csv(
  file = "../../data/clean/cleaned_pavlovia.csv"
)
# Load fitted models
load("../../data/temp/models/study_1_fitted.RData")
```

@@

```{r}
#| label: fig-hook-1
#| fig-cap: Probability of voting for candidate with red yard sign versus white
#| 
red <- predicted_prob_bar(
  fitted_model = list_fitted[["h_3"]][["red"]]
  , x_axis = "trial_1_red_stimuli"
  , treatment = "Red"
  , hypothesis = "H3"
  , x_label = "Party ID"
  , y_label = "Pr(Vote for candidate)"
  , legend_title = "Treatment"
)

```{r}
#| label: fig-hook-2
#| fig-cap: Probability of voting for candidate with blue yard sign versus white
blue <- predicted_prob_bar(
  fitted_model = list_fitted[["h_3"]][["blue"]]
  , x_axis = "trial_1_blue_stimuli"
  , treatment = "Blue"
  , hypothesis = "H3"
  , x_label = "Party ID"
  , y_label = "Pr(Vote for candidate)"
  , legend_title = "Treatment"
)
blue
```

@@

## Question guiding the project

@@

## De-mystifying some of the jargon I might use

@@

- Object
  - refers to anything out in the world that we are interacting with
  - here, these objects are political
- Affect $\neq$ emotion
  - refers to the larger state of being
  - emotion refers to consciously labeled reactions
  - valance refers to pre-consciously 
- "Hot Cognition"
  - refers to the idea that some memories and experiences are more accessible or come to the top of the mind quicker.
  - this coincides with memories and experiences that are affect-laden

## Dominant models of information processing

### Borrowing concepts from social neuroscience and psychology

@@

- Memories:
  -  are multi-dimensional encodings of previous experiences [@fazio_attitudes_2007]
    - e.g., remember giving your first job talk
      - the layout of the room
      - the colors of the chairs
      - what the room smelled like
      - what ambient sounds were there
      - how you felt
  - affectively laden [@fazio_attitudes_2007;@winkielman_emotion_2011; @kensinger_affective_2022]
  - easier retrieved when recent and/or shares multiple commonalities with current object

### Zaller's [-@zaller_nature_1992] Memory model

@@

```{mermaid}
%%| label: fig-zaller
%%| fig-cap: Memory model of info processing
graph TD;
A[Sampled pre-existing info] -->|Stronger impacts when less pre-existing knowledge| C[Updated attitudes]
B[New info] --> C
```

@@

```{r}
#| label: sim-visualization

df <- fabricatr::fabricate(
  N = 10000,
  pop = rnorm(n = N, mean = 2, sd = 1),
  sample_strong = rnorm(n = N, mean = 2, sd = 0.5),
  updated_strong = rnorm(n = N, mean = 1.9, sd = 1),
  sample_weak = rnorm(n = N, mean = 2, sd = 2),
  updated_weak = rnorm(n = N, mean = 1.5, sd = 1)
)
```

```{r}
#| label: fig-pre-existing-info
#| fig-cap: Set of pre-existing attitudes
ggplot(data = df) +
  geom_density(aes(x = pop)) +
  geom_vline(aes(xintercept = 2), linetype = 1) +
  theme_minimal()
```

@@

```{r}
#| label: fig-sample-pre-existing-info
#| fig-cap: Pre-existing attitudes used
#| fig-subcap:
#|   - Unknowledgeable person
#|   - Knowledgeable person
ggplot(data = df) +
  geom_density(aes(x = sample_weak)) +
  geom_vline(aes(xintercept =  2), linetype = 1) +
  theme_minimal()
ggplot(data = df) +
  geom_density(aes(x = sample_strong)) +
  geom_vline(aes(xintercept =  2), linetype = 1) +
  theme_minimal()
```

@@

```{r}
#| label: fig-new-attitude
#| fig-cap: Updated set of attitudes
#| fig-subcap:
#|   - Unknowledgeable person
#|   - Knowledgeable person
ggplot(data = df) +
  geom_density(aes(x = updated_weak)) +
  geom_vline(aes(xintercept =  2), linetype = 1) +
  geom_vline(aes(xintercept =  1.5), linetype = 2) +
  theme_minimal()
ggplot(data = df) +
  geom_density(aes(x = updated_strong)) +
  geom_vline(aes(xintercept =  2), linetype = 1) +
  geom_vline(aes(xintercept =  1.9), linetype = 2) +
  theme_minimal()
```

### Lodge and Taber's [-@lodge_rationalizing_2013] Online model

@@

```{mermaid}
%%| label: fig-online-model
%%| fig-cap: Online model of info processing
%%| fig-width: 6.5
graph TD;
A[Pre-existing beliefs] --> C[Updated beliefs]
B[New information] -->|If it aligns with current beliefs| C
```

## Amending those models with the Snap-judgment model

>>>>

```{r}
#| label: setup-block
# Set seed
set.seed(12102023)
# Load libraries
box::use(
  ggplot2[
    ggplot
    , aes
    , geom_density
    , theme_minimal
    , geom_vline
    , labs
    , scale_y_discrete
    , scale_x_discrete
  ]
  , ggdist[
    stat_halfeye
  ]
  , modelsummary[
    datasummary_skim
    , datasummary_balance
    , modelsummary
  ]
  , marginaleffects[
    avg_slopes
    , posterior_draws
  ]
  , ../R/predicted_prob_bar[...]
  , data.table[...]
)

# Load cleaned data
list_df <- list()
list_df[["cleaned"]] <- read.csv(
  file = "../../data/clean/cleaned_pavlovia.csv"
)
# Load fitted models
  #* Pre-test
load(
    file="../../data/models/prr/pre-test/pre-test_party_model.RDS"
  )
  #* Main studies
load("../../data/temp/models/study_1_fitted.RData")
```

@@

```{r}
#| label: fig-hook-1
#| fig-cap: Perceptions of candidate's party affiliation based on color of yard sign

partyAMES <- avg_slopes(
  PartyModel,
  type = "link"
) |>
posterior_draws()

ggplot(
  data = partyAMES,
  aes(
    x = draw,
    y = term
  )
) +
stat_halfeye(
  slab_alpha = 0.45
) +
theme_minimal() +
labs(
  x = "Average Marginal Effect",
  y = "",
  caption = "Data source: Pre-test.\n Distribution of AME's for model posterior draws using HDI's.\n Interpreted as the logged odds of labelling the candidate as a partisan\n given whether they viewed a red yard sign or blue yard sign\n while holding all other variables at their mean value.\n Dot reflects median posterior draw where bars reflect 90% and 95% credible intervals."
) +
scale_y_discrete(
  labels = c(
    "Blue Treatment",
    "Red Treatment"
  )
) +
ggplot2::scale_x_continuous(
  breaks = c(-3, 0, 3)
  , labels = c(
    "Strong Democrat", "Independent", "Strong Republican"
  )
)
```

@@

# Question guiding the project

@@

# What do we know so far?

## The study of visuals in politics

@@

![[@bucy_editors_2021]](../../assets/slides/imgs/bucy_joo.png)

@@

![](../../assets/slides/imgs/dan_arendt.png){.absolute bottom=0 left=0}
![](../../assets/slides/imgs/lindholm_et-al.png){.absolute top=200 left=0}
![](../../assets/slides/imgs/nina_et-al.png){.absolute top=0 left=0}

@@

![[@schloss_politics_2014]](../../assets/slides/imgs/schloss_palmer.png){.absolute top=0 height=400}
![[@williams_whats_2022]](../../assets/slides/imgs/williams_et-al.png){.absolute bottom=200 right=0 height=400}
![[@casiraghi_colors_2022]](../../assets/slides/imgs/casiraghi_et-al.png){.absolute bottom=0 right=0 height=400}

## Canonical models of **political** information processing

@@

## Integrating key perspectives from the cognitive sciences

@@

![[@collins_spreading-activation_1975]](../../assets/slides/imgs/collins_loftus.png){.absolute top=100}
![[@fazio_attitudes_2007]](../../assets/slides/imgs/fazio.png){.absolute bottom=100}

@@

![[@sander_cambridge_2013]](../../assets/slides/imgs/affective_neuro.jpg){.absolute top=100 height=600}
![[@adolphs_emotion_2011]](../../assets/slides/imgs/affective_neuroscience.webp){.absolute bottom=100 right=0}

@@

![[@zajonc_feeling_1980]](../../assets/slides/imgs/zajonc.png){.absolute top=50 right=50 height=300}
![](../../assets/slides/imgs/erisen_et-al.png){.absolute bottom=50 left=50}

## Canonical models of **political** information processing

![[@zaller_nature_1992]](../../assets/slides/imgs/zaller.jpg){.absolute left=0}
![](../../assets/slides/imgs/zaller_feldman.png){.absolute right=0}
![[@taber_motivated_2006]](../../assets/slides/imgs/taber_lodge.png){.absolute top=200 right=0}
![[@lodge_rationalizing_2013]](../../assets/slides/imgs/lodge_taber.jpg){.absolute top=100}
![[@taber_illusion_2016]](../../assets/slides/imgs/taber_lodge_2016.png){.absolute top=400}

## A snap-judgment model of **visual** political information processing
